---
title: "Reproducible Project"
author: "Joel Samson"
date: "February 26, 2018"
output: 
  html_document: 
    keep_md: yes
---   

### Download data, unzip, read 
```{r, cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, warning=FALSE}

tmp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", tmp)
unzip(tmp, "activity.csv")
Newactivity <- read.table("activity.csv", sep=",", header=T)
```

### Take a copy of data and format the field date as date format 
```{r, cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, warning=FALSE}
activityCopy <- Newactivity
str(Newactivity$date)
Newactivity$Date <- as.Date(Newactivity$date, "%Y-%m-%d")
Newactivity$interval <- as.factor(Newactivity$interval)
l <- levels(Newactivity$interval)
```

### 1) What is mean total number of steps taken per day?

Total number of steps taken each day
```{r, echo=FALSE}
totalSteps <- aggregate(steps ~ date, data = Newactivity, sum, na.rm = TRUE)
totalSteps
```

Mean and Median of steps taken each day
```{r, echo=FALSE}
rmean <- mean(totalSteps$steps)
rmean
rmedian <- median(totalSteps$steps)
rmedian
```

### Draw the histograms for Total Number of Steps By Each Day
```{r, echo=FALSE, Error=FALSE, cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
totalSteps <- tapply(Newactivity$steps, Newactivity$date, sum, na.rm=TRUE)
#par(mfrow=c(2,1))
hist(totalSteps, breaks=10, col="green", main="Total Number of steps each day", xlab="Total Number of Steps")
```

### 2) What is the average daily activity pattern?
1.Make a time series plot (i.e. ) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=FALSE, cache=FALSE, results=FALSE, warning=FALSE}
steps_by_interval <- aggregate(steps ~ interval, Newactivity, mean)
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Average Number of Steps",main="Average Number of Steps per Day by Interval")
```

2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r,echo=FALSE, Error=FALSE, cache=FALSE}
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval
```

### 3) Imputing missing values
Total Number of Missing values: 
```{r, echo=FALSE}
NATotal <- sum(!complete.cases(Newactivity))
NATotal
```

Summary of Before and After Imputing missing values
```{r, echo=FALSE}
missing <- is.na(Newactivity)
activityCopy$steps[is.na(activityCopy$steps)] <- tapply(activityCopy$steps, activityCopy$interval, mean, na.rm = TRUE)
totStepsImp <- tapply(activityCopy$steps, activityCopy$date, sum)

summary(totalSteps)
summary(totStepsImp)
```

```{r, echo=FALSE}
par(mfrow=c(2,1))
hist(totalSteps, col="red", xlab="Average Total Number of Steps Before Imputing", main="", breaks=10)
title("Distribution of the total Number of steps each day Before and After Imputing")
hist(totStepsImp, col="cyan", main="", xlab="Average Total Number of Steps After Imputing", breaks=10)
```

### 4) Are there differences in activity patterns between weekdays and weekends?
```{r, echo=FALSE}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
activityCopy$dow = as.factor(ifelse(is.element(weekdays(as.Date(activityCopy$date)),weekdays), "Weekday", "Weekend"))
StepsTotalUnion <- aggregate(steps ~ interval + dow, activityCopy, mean)
library(lattice)
xyplot(StepsTotalUnion$steps ~ StepsTotalUnion$interval|StepsTotalUnion$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```